---
title: "Brownlows Bets"
author: "Laizha Grajeda De Leon"
date:   "Summer 2022"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
   
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.

## what is their average earnings

```{r}
mean(Substitute)
```
[1] 279.175


The average earnings would be about $280 if we were to hire 40 Brownlows to do the bet.


## the standard deviation and a 95% confidence interval for their earnings?
  
```{r}
sd(Substitute)
```
[1] 34.75425


The standard deviation is 34.574, and based on the central limit theorm, the 95% interval will be within two standard deviation values above and below the mean. So, if we go 68 above and below 280, there will be 95% of the data in the distribution. 
  
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->


# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 


```{r}
Fair <- numeric (length=40
for (j in 1:1000) {
   Fair[j] =  BrLw(odds=4.38)
}
```





```{r}
hist(Fair)
```



We are going to hire 100 Brownlows to repeat what Brownlow did

```{r}
Substitute <- numeric (length=1000)
for (j in 1:1000) {
   Substitute[j] =  BrLw(odds=4.38)
}
```


```{r}
hist(Substitute)
```

```{r}
mean(Substitute)
```


```{r}
sd(Substitute)
```

[1] 53.70783




When the odds are 4.38:1, Brownlow profits centers around $0 as show in the hitsogram and therefore, the fair odds is when Bownlow wager is $4.38 for every $1 that the students would wager for th bet. 



Using the mean and standard deviation (subtracting 110 from the mean and adding 110 to the mean) we found the interval we are looking for, and it looks like the profit is centered around zero. If the profit is centered around zero, that is fair odds. More specifically, the fair odds is when Brownlow would wager 4.38 dollars for every dollar the students would wager.


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->




# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->

```{r}
Sub <- numeric(length = 1000)

for (j in 1:1000) {
  Sub[j] = BrLw(numClass = 29)
}
```

```{r}
hist(Sub)
```

With a 2:1 odds, it appears that there must be about 28 or 29 people in the class in order for the bet to have a fair outcome or, in other words, have the profit be centered towards zero. the distribution when there are 29 students in the class,the histogram above shows and given than the peak lies on the right of zero,Brownlow still gains money from the bet. Meanwhile, making the class size 28 produces a distribution in which the peak lies to the left of zero which means that Brownlow loses money in the bet. Therefore, we can assume that the class size must be between 28 and 29 for there to be a fair bet with a 2:1 odds.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

We saw that in this in this write-up, we saw how to create a function that would simulate bets between Brownlow and students in a class. By manipulating this function, we are able to figure out the number of the students who were playing the bets in order to have a profit of $0, as well as the fair odds. Furthermore, we applied the central limit theorem to find the 95% confidence interval if the bets are played with fair odds. 



